Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-08-20 prop_letter.tex
prop_letter.tex Updated by Gerardo...

\begin{thm}\label{thm:Shapley}
  For QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$, if a subspace
  of events~$\events_C\subseteq\events$ commutes, then
  $\coreBorn\left(\bar{\mu},\events_C\right)\ne\emptyset$.
\end{thm}

@@

In conventional quantum mechanics the expectation value of an
observable is defined in Eq.~\ref{eq:quantum-expectation} as a precise
real number. The generalization of quantum measures to QIVPMs implies
that the real-valued expectation value will be generalized to an
interval.

@@

\paragraph*{$\delta$-Determinism.} The first step in our formalization
is to introduce a family of QIVPMs parameterized by an uncertainty
$\delta$ which we call $\delta$-deterministic QIVPMs.

\begin{definition}\label{def:delta-deterministic} A
  QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$ is
  $\delta$-deterministic if, for every event $P\in\events$, we have
  that either 
  $\bar{\mu}\left(P\right)\subseteq\left[0,\delta\right]$ or
  $\bar{\mu}\left(P\right)\subseteq\left[1-\delta,1\right]$. 
\end{definition}

@@

\begin{thm}
\label{thm:Kochen-Specker} {\rm [0-deterministic variant of Kochen-Specker]}
Given a Hilbert space $\Hilb$ of dimension~$d\ge3$, there is no
0-deterministic
measure~$\mu^{\mathrm{D}}:\events\rightarrow\left\{ 0,1\right\}$.
\end{thm}

To prove this result, we proceed by assuming a 0-deterministic QIVPM
$\bar{\mu}$ and derive a contradiction using the same nine observables
$\mathbf{O}_{ij}$ used in the ``Mermin-Peres'' proof of the
Kochen-Specker theorem~\cite{Mermin1990Simple,peres1995quantum}:
{\renewcommand{\arraystretch}{2}%
\begin{center}
\begin{tabular}{|@{\quad}c@{\quad}|@{\quad}c@{\quad}|@{\quad}c@{\quad}|}
\hline 
$\mathbb{1}\otimes\sigma_{z}$  & $\sigma_{z}\otimes\mathbb{1}$  & $\sigma_{z}\otimes\sigma_{z}$ \tabularnewline
\hline 
$\sigma_{x}\otimes\mathbb{1}$  & $\mathbb{1}\otimes\sigma_{x}$  & $\sigma_{x}\otimes\sigma_{x}$ \tabularnewline
\hline 
$\sigma_{x}\otimes\sigma_{z}$  & $\sigma_{z}\otimes\sigma_{x}$  & $\sigma_{y}\otimes\sigma_{y}$ \tabularnewline
\hline 
\end{tabular}\,,
\par\end{center}
}
\noindent where $\left\{\mathbb{1},\sigma_{x},\sigma_{y},\sigma_{z}\right\}$ are
Pauli matrices acting on a pair of spin-$\frac{1}{2}$ states. We now define the
expectation value of the QIVPM for our Mermin-Peres square.

Since the expectation value~$\expval{\cdot}_{\bar{\mu}}$ satisfies
these two properties, the observables in the Mermin-Peres square must
have the expectation value~$\left\{ 1\right\} $ or
$\left\{ -1\right\} $.  In each row and column, the operators commute,
and each operator is the product of two others, except in the third
column, where
$\left(\sigma_{z}\otimes\sigma_{z}\right)\left(\sigma_{x}\otimes\sigma_{x}\right)=-\sigma_{y}\otimes\sigma_{y}$.
Therefore, it is clearly impossible to find a QIVPM~$\bar{\mu}$ such
that $\langle \mathbf{O}_{ij}\rangle_{\bar \mu}$ satisfies
Eq.~(\ref{eq:Kochen-Specker-Rules}).

@@

\begin{thm}\label{cor:Kochen-Specker-IVPM} Given a Hilbert space $\Hilb$
  of dimension~$d\ge3$, there is no $\delta$-deterministic QIVPM for
  $\delta<\frac{1}{3}$.\end{thm}

\begin{proof} Suppose there is a $\delta$-deterministic
  QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$, we claim that the
  following function~$\bar{\mu}^{\textrm{D}}:\events\rightarrow\left\{
    \imposs,\necess\right\} $ is a cryptodeterministic QIVPM:
\begin{equation}
\bar{\mu}^{\textrm{D}}\left(P\right)=\begin{cases}
\imposs\,, & \textrm{ if }\bar{\mu}\left(P\right)\subseteq\left[0,\delta\right]\:;\\
\necess\,, & \textrm{ if }\bar{\mu}\left(P\right)\subseteq\left[1-\delta,1\right]\:.
\end{cases}
\end{equation}
However, since we know that there is no cryptodeterministic QIVPM,
$\bar{\mu}^{\textrm{D}}$ cannot exist and hence $\bar{\mu}$ cannot
exist.

@@

When $\bar{\mu}^{\textrm{D}}\left(P_{0}\right)=\bar{\mu}^{\textrm{D}}\left(P_{1}\right)=\imposs$,
we have both $\bar{\mu}\left(P_{0}\right)$ and $\bar{\mu}\left(P_{1}\right)\subseteq\left[0,\delta\right]$
which implies $\bar{\mu}\left(P_{0}+P_{1}\right)\subseteq\left[0,2\delta\right]$.
Since $\delta<\frac{1}{3}\Rightarrow2\delta<1-\delta$, $\bar{\mu}\left(P_{0}+P_{1}\right)$
cannot be a subset of $\left[1-\delta,1\right]$, but $\bar{\mu}\left(P_{0}+P_{1}\right)$
is a subset of either $\left[0,\delta\right]$ or $\left[1-\delta,1\right]$.
Hence, $\bar{\mu}\left(P_{0}+P_{1}\right)$ must be a subset of $\left[0,\delta\right]$,
which implies $\bar{\mu}^{\textrm{D}}\left(P_{0}+P_{1}\right)=\imposs$. Thus the
theorem holds when $\delta<\frac{1}{3}$.
\end{proof}

@@

A $\frac{1}{3}$-deterministic QIVPM can be given by
$\bar{\mu}\left(P\right)=\left[x/3,x/3\right]$, where $P$ projects
onto an $x$-dimensional subspace of a three-dimensional Hilbert space.

>>>>

\begin{thm}[Non-empty Core for Compatible Measurements] \label{thm:Shapley}
  For QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$, if a subspace
  of events~$\events_C\subseteq\events$ commutes, then
  $\coreBorn\left(\bar{\mu},\events_C\right)\ne\emptyset$.
\end{thm}

@@

In conventional quantum mechanics the expectation value of an
observable is defined in Eq.~\eqref{eq:quantum-expectation} as a precise
real number. The generalization of quantum measures to QIVPMs implies
that the real-valued expectation value will be generalized to an
interval.

@@

%\paragraph*{$\delta$-Determinism.}

The first step in our formalization
is to introduce a family of QIVPMs parameterized by an uncertainty
$\delta$ which we call $\delta$-deterministic QIVPMs.

\begin{definition}[$\delta$-Determinism]\label{def:delta-deterministic} A
  QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$ is
  $\delta$-deterministic if, for every event $P\in\events$, we have
  that either 
  $\bar{\mu}\left(P\right)\subseteq\left[0,\delta\right]$ or
  $\bar{\mu}\left(P\right)\subseteq\left[1-\delta,1\right]$. 
\end{definition}

@@

\begin{thm}
[0-Deterministic Variant of Kochen-Specker] \label{thm:Kochen-Specker} 
Given a Hilbert space $\Hilb$ of dimension~$d\ge3$, there is no
0-deterministic
measure~$\mu^{\mathrm{D}}:\events\rightarrow\left\{ 0,1\right\}$.
\end{thm}

To prove this result, we proceed by assuming a 0-deterministic QIVPM
$\bar{\mu}$ and derive a contradiction using the same nine observables
$\mathbf{O}_{ij}$ used in the ``Mermin-Peres'' proof of the
Kochen-Specker theorem~\cite{Mermin1990Simple,peres1995quantum}:
{\renewcommand{\arraystretch}{2}%
\begin{center}
\begin{tabular}{|@{\quad}c@{\quad}|@{\quad}c@{\quad}|@{\quad}c@{\quad}|}
\hline 
$\mathbb{1}\otimes\sigma_{z}$  & $\sigma_{z}\otimes\mathbb{1}$  & $\sigma_{z}\otimes\sigma_{z}$ \tabularnewline
\hline 
$\sigma_{x}\otimes\mathbb{1}$  & $\mathbb{1}\otimes\sigma_{x}$  & $\sigma_{x}\otimes\sigma_{x}$ \tabularnewline
\hline 
$\sigma_{x}\otimes\sigma_{z}$  & $\sigma_{z}\otimes\sigma_{x}$  & $\sigma_{y}\otimes\sigma_{y}$ \tabularnewline
\hline 
\end{tabular}\,,
\par\end{center}
}
\noindent where $\left\{\mathbb{1},\sigma_{x},\sigma_{y},\sigma_{z}\right\}$ are
Pauli matrices acting on a pair of spin-$\frac{1}{2}$ states. We now define the
expectation value over the QIVPM for our Mermin-Peres square.

Since the expectation value~$\expval{\mathbf{O}_{ij}}_{\bar{\mu}}$ satisfies
these two properties, the observables in the Mermin-Peres square must
have the expectation value~$\left\{ 1\right\} $ or
$\left\{ -1\right\} $.  In each row and column, the operators commute,
and each operator is the product of two others, except in the third
column, where
$\left(\sigma_{z}\otimes\sigma_{z}\right)\left(\sigma_{x}\otimes\sigma_{x}\right)=-\sigma_{y}\otimes\sigma_{y}$.
Therefore, it is clearly impossible to find a QIVPM~$\bar{\mu}$ such
that $\langle \mathbf{O}_{ij}\rangle_{\bar \mu}$ satisfies
Eq.~(\ref{eq:Kochen-Specker-Rules}).

@@

\begin{thm}[Finite-precision Extension of Kochen-Specker] \label{cor:Kochen-Specker-IVPM} Given a Hilbert space $\Hilb$
  of dimension~$d\ge3$, there is no $\delta$-deterministic QIVPM for
  $\delta<\frac{1}{3}$.\end{thm}

%\begin{proof} 
Suppose there is a $\delta$-deterministic
  QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$, we claim that the
  following function~$\bar{\mu}^{\textrm{D}}:\events\rightarrow\left\{
    \imposs,\necess\right\} $ is a cryptodeterministic QIVPM:
\begin{equation}
\bar{\mu}^{\textrm{D}}\left(P\right)=\begin{cases}
\imposs\,, & \textrm{ if }\bar{\mu}\left(P\right)\subseteq\left[0,\delta\right]\:;\\
\necess\,, & \textrm{ if }\bar{\mu}\left(P\right)\subseteq\left[1-\delta,1\right]\:.
\end{cases}
\end{equation}
However, since we know that there is no cryptodeterministic QIVPM,
$\bar{\mu}^{\textrm{D}}$ cannot exist and hence $\bar{\mu}$ cannot
exist.

@@

When $\bar{\mu}^{\textrm{D}}\left(P_{0}\right)=\bar{\mu}^{\textrm{D}}\left(P_{1}\right)=\imposs$,
we have both $\bar{\mu}\left(P_{0}\right)$ and $\bar{\mu}\left(P_{1}\right)\subseteq\left[0,\delta\right]$
which implies $\bar{\mu}\left(P_{0}+P_{1}\right)\subseteq\left[0,2\delta\right]$.
Since $\delta<\frac{1}{3}\Rightarrow2\delta<1-\delta$, $\bar{\mu}\left(P_{0}+P_{1}\right)$
cannot be a subset of $\left[1-\delta,1\right]$, but $\bar{\mu}\left(P_{0}+P_{1}\right)$
is a subset of either $\left[0,\delta\right]$ or $\left[1-\delta,1\right]$.
Hence, $\bar{\mu}\left(P_{0}+P_{1}\right)$ must be a subset of $\left[0,\delta\right]$,
which implies $\bar{\mu}^{\textrm{D}}\left(P_{0}+P_{1}\right)=\imposs$. Thus the
theorem holds when $\delta<\frac{1}{3}$.
%\end{proof}

@@

A $\frac{1}{3}$-deterministic QIVPM can be given by
$\bar{\mu}\left(P\right)=\left[n/3,n/3\right]$, where $P$ projects
onto an $n$-dimensional subspace of a three-dimensional Hilbert space.
